Adding selection to \stdpopsim: oh how the tables have turned

### 2024-12-18 analysis2_manuscript.tex
Edits from Ryan, mostly on DFE section.

% add commands for the DFE inference methods
\newcommand{\dfe}{\texttt{dfe-alpha}\xspace}
\newcommand{\polydfe}{\texttt{polyDFE}\xspace}
\newcommand{\dadi}{$\partial$\texttt{a}$\partial$\texttt{i}\xspace}
\newcommand{\grapes}{\texttt{GRAPES}\xspace}

@@

@@

% \author[1,+]{M. Elise Lauterbur}
% \author[3,*]{Ariella L. Gladstein}
 \author[4,*]{Graham Gower}
 \author[5*]{Nathaniel S. Pope}
 \author[5*]{Murillo F. Rodrigues}
 \author[5*]{Silas Tittes}
\author[6]{Georgia Tsambos}
\author[34]{Linh N. Tran}
\author[2]{Maria Izabel A. Cavassim}
% \author[5,7]{Jeff Adrion}
% \author[5]{Saurabh Belsare}
% \author[8]{Arjun Biddanda}
% \author[5]{Victoria Caudill}
% \author[9]{Jean Cury}
% \author[10]{Ignacio Echevarria}
% \author[11]{Benjamin C. Haller}
% \author[12,13]{Ahmed R. Hasan}
\author[14,15]{Xin Huang}
% \author[16]{Leonardo Nicola Martin Iasi}
% \author[17]{Ekaterina Noskova}
% \author[18]{Jana Obšteter}
% \author[19]{Vitor Antonio Corrêa Pavinato}
% \author[20,21]{Alice Pearson}
% \author[22,23]{David Peede}
% \author[24]{Manolo F. Perez}
\author[5]{Chris C. R. Smith}
% \author[25]{Jeffrey P. Spence}
% \author[5]{Anastasia Teterina}

@@

% \affil[14]{\small{Department of Evolutionary Anthropology, University of Vienna, Vienna, Austria}}
% \affil[15]{\small{Human Evolution and Archaeological Sciences (HEAS), University of Vienna, Vienna, Austria}}

@@

% \affil[34]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson AZ 85721, USA}}

@@

    % simulation
    To meet the growing need for the interpretation,
    analysis, and exploration of realistic and complex evolutionary models
    the field of population genetics has increasingly turned to simulation.
    Simulation in population genetics has a long history 
    including both backward-in-time coalescent simulations
    \citep{kingman1982genealogy,hudson1983testing, hudson1990gene}
    and forward-in-time simulations of complex demography and selection
    \citep[e.g.,][]{gillespie1984molecular,thornton2014c++, haller2019slim}.
    The development of simulation tools has been driven by the need to
    understand the effects of complex evolutionary processes on genetic
    variation \citep[e.g.,][]{galloway2020few}, to provide a null model for hypothesis testing
    \citep[e.g.,][]{hudson1992statistical,hudson1994evidence,sabeti2002detecting},
    to explore the power and limitations of statistical methods \citep[e.g.,][]{przeworski2002signature},
    and increasingly to provide a basis for machine learning and other
    simulation-based inference methods \citep[e.g.,][]{beaumont2002approximate,pavlidis2010searching,lin2011distinguishing,kern2018diplos,mughal2019localizing,sanchez2021deep,wang2021automatic}.
    While this is so, joint simulation of complex demography and selection
    is challenging, and requires a deep understanding of the underlying
    evolutionary processes, and necessitates a number of parameter choices including
    the strength of selection and the
    recombination rate, as well as a parameterized demographic model.
    This is a daunting task for many researchers, and can be a barrier to
    the adoption of simulation-based methods in population genetics.
    Furthermore, different simulation and inference tools may use different structures
    and parameter scalings for their models.

    % stdpopsim
    In light of the complexities associated with estimating and simulating
    population genetic models, it is no surprise that
    a lingering challenge with simulation in population genetics has been
    reproducibility and the ability to share and compare results among 
    researchers \citep[e.g.,][]{ragsdale2020lessons}.
    This challenge has been addressed in part by the development
    of community resources for sharing and distributing simulation software
    via the \stdpopsim project \citep{adrion2020community}. \stdpopsim
    provides a standardized interface for accessing a wide range of
    population genetic models, and has begun to be be widely adopted by the community. %maybe add citations here?
    While that is so, the original version of \stdpopsim did not include
    models of selection, which is a major limitation for more empirical
    applications of population genetic simulation. In particular, modeling selection
    through simulation is critical for understanding processes such
    as adaptation, the effects of selective sweeps \citep[e.g.][]{braverman1995hitchhiking,fay2000hitchhiking,przeworski2002signature,przeworski2005signature,schrider2015soft}, and the impact of
    background selection on genetic diversity
    \citep[e.g.][]{charlesworth1993effect,charlesworth1995pattern,williamson2002genealogy,ewing2016consequences,torres2020temporal}.
    Ideally, one would like
    to have a single, unified framework for simulating both neutral and
    non-neutral evolutionary processes, and to be able to compare the
    results of these simulations to empirical data in a manner that is
    both accessible to a wide range of researchers and highly reproducible. 
    Further the framework should include the complex realities of 
    genomes, including heterogenous recombination rates, 
    variation in the size and density of functional elements, and
    non-equilibrium demographic histories. Ideally, one should be able
    to model all of these features after estimates/annotations obtained
    from a species of interest \citep[e.g.][]{schrider2020background,rodrigues2024shared}.

@@

    \begin{wrapfigure}{l}{0.5\textwidth}
        \vspace{-0.0cm}
        \includegraphics[width=\linewidth]{figures/schematics/catalog.pdf}
        \caption{\label{fig:schematic}
        \textbf{(A)} A schematic of the \stdpopsim catalog and the different components that
        can be used in simulations. 
        \textbf{(B)} The python API and \textbf{(C)} the command line interface.
        The user specifies a species, a portion of
        the genome to simulate, and optionally a genetic map, a model of
        demography, and a model of selection that itself is composed of a
        distribution of fitness effects (DFE) and a set of functional
        annotations.}
    \end{wrapfigure}
    \hfill

@@

    The vast majority of published DFE estimates are based on a limited number of species,
    so \stdpopsim currently has implemented published DFEs from
    four species: \textit{Arabidopsis thaleana} \citep{huber2018gene}, \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining},
    humans \citep{huber2017determining,kim2017inference}, and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
    While that is so, these DFEs can be applied to other species
    in the catalog. For instance, one could simulate a butterfly species with a gamma DFE originally
    estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
    Furthermore, the user can specify a custom DFE, and provide their own annotations
    of functional elements to simulate selection in a species for which we do not yet have 
    a published DFE included in the catalog. This flexibility allows for a wide range of
    models of selection to be simulated. %, and for the results of these simulations to be
    %compared to empirical data or used as a null model for hypothesis testing.
    % DRS: commenting that stuff out because there are obviously other use cases that could be
    % listed here (e.g. ML/ABC, and if we don't know enough about a species to do that then
    % we probably have no business doing hypothesis testing either, which could be just as
    % misleading or maybe even more so), so the simplest solution is to cut this as it is
    % not needed here.

@@

    Using \stdpopsim, we can easily simulate genetic data under a model of selection and then 
    examine the performance of different methods for inferring $N_e(t)$ from these simulated 
    data. In Figure \ref{fig:1pop-human-demography}, we show the results of such a benchmark
    where we have simulated genomes under a model of human out of Africa (OOA) demography
    with and without purifying selection acting on exons.
    For this simulation we have used
    a genetic map from the HapMap Project \citep{international2007second} along with a
    DFE for deleterious alleles (estimated by \cite{huber2017determining})
    that acts on exonic positions defined from the HAVANA group release 104 for the human genome.
    % DRS: Maybe we should say something about how easy it was to do these simulations, and maybe even show
    % some example code somewhere to prove our point? Could help emphasize the utility of this resource.
    These simulations were then used to estimate $N_e(t)$ using four methods: 1) \msmc \citep{Schiffels2020}, 
    2)\stairway method \citep{liu2020stairway}, 3) \gone \citep{santiago2020recent}, and 4) \smcpp \citep{terhorst2017robust}.
    In this and what follows we performed 3 replicate whole genome simulations, so uncertainty in estimates might
    be roughly assessed, however we are mindful of wasted computational resources as thorough benchmarking
    is not the main purpose of this manuscript.

    In Figure \ref{fig:1pop-human-demography}, we show six panels, each of which shows the estimated $N_e(t)$
    from data simulated under the five-population model of human out-of-Africa demography
    from Ragsdale \textit{et al.} (2019)
    with samples taken from three modern geographic groups (CEU, CHB, and YRI from left to right).
    The left column shows the results of $N_e(t)$ estimates made from simulations without selection, 
    while the right column shows the results
    from simulations with deleterious mutations occurring in exons.
    In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$
    in various colors for the four methods. Dotted lines on the top row are to visually reinforce that inference is done
    from a model without selection. We see that these methods generally perform reasonably well in the absence of selection, 
    although GONE seems to overestimate $N_e(t)$ for the CHB sample and to a lesser extent in the CEU sample.
    Presumably this could be due to patterns of linkage disequilibrium that result from the combination
    of bottleneck and growth in these populations.
    In the presence of selection, different methods behave differently. 
    For instance stairwayplot, which uses the SFS to infer $N_e(t)$, 
    is reasonably close to the true $N_e(t)$ both with and without selection,
    although it seems to infer significant population size changes that are not part of the true history.
    Likewise \smcpp seems to be robust to the presence of selection,
    but the timing of population size changes is slightly inaccurate. 
    \msmc seems to be the most affected by the presence of selection, 
    estimates of $N_e(t)$ varying quite a bit from the true $N_e(t)$ particularly in the very recent past,
    although estimates from \msmc have quite a bit of added noise even in the absence of selection.

@@

    \begin{figure}[t]
        \centering
        \includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
        \caption{
        \label{fig:1pop-human-demography}
        Performance of methods to infer $N_e(t)$ from a human out-of-Africa model \citep{ragsdale2019models}
        with and without purifying selection on exons. The left column shows estimates of $N_e(t)$ from simulations
        without selection, while the right column shows estimates of $N_e(t)$ from simulations with a gamma-distributed   
        DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
        \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.  
        }
    \end{figure}
    %DRS-fig: reminder to update the legend to just say ``exons'' instead of ensembl_havana_whatever...

@@

\section*{Estimation of the Distribution of Fitness effects}
    \label{dfe}
    Another common application of population genetic inference is to estimate the distribution of fitness effects (DFE)
    from genetic data. Here, we show how the new models of selection in \stdpopsim can be used to benchmark and compare
    the performance of different methods for inferring the DFE. We performed simulations using the 
    human and vaquita porpoise models described above, with complex demographic histories and models of selection acting on exons.

    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_OutOfAfricaArchaicAdmixture_5R19_Gamma_K17_ensembl_havana_104_exons_DFE_plot}
        \caption{A comparison of methods for inferring the distribution of fitness effects (DFE) from genetic data.
        Simulations were performed using a human out-of-Africa model \citep{ragsdale2019models} with a gamma-distributed DFE
        acting on exons parameterized by a mean selection coefficient and a shape parameter. Estimates of the 
        parameters of the DFE are shown in the left ($\lvert E(s) \rvert $) and right hand panels (shape) respectively.
        The human model has three extant populations (CEU, CHB, and YRI), and parameter estimates from each
        sample are shown in different colors.
        Estimates from three different methods are shown: 1) \grapes \cite{galtier2016adaptive}, 2) \polydfe \citep{tataru2020polydfe},
        and 3) \dadi \citep{gutenkunst2009inferring}.}
        \label{fig:homsap-dfe.ooa}
    \end{figure}

    In Figures \ref{fig:homsap-dfe.constant} and \ref{fig:homsap-dfe.ooa} we show a comparison of estimates
    from three software packages for inferring the DFE from genetic data:
    \dadi \citep{gutenkunst2009inferring}, \polydfe \citep{tataru2020polydfe}, 
    and \grapes \citep{galtier2016adaptive}.
    In each case we note that the DFE is estimated from segregating sites only,
    without the use of substitution data, as a current limitation of \stdpopsim
    is that we are not able to simulate an outgroup sequence to estimate
    divergence. 
    Figure \ref{fig:homsap-dfe.constant} shows estimates from a single, constant size population of humans, 
    while Figure \ref{fig:homsap-dfe.ooa} shows estimates from a model of human out-of-Africa demography.
    In a constant size population, all methods perform reasonably well, although \dadi seems to
    perform slightly worse than the other methods examined, at least for the small number of replicates we simulated here.
    Notably only \polydfe overlapped the true mean selection coefficient, whereas 
    the other methods do not. 
    For the shape parameter in constant size simulations \grapes and \polydfe are roughly
    equivalent in their performance whereas \dadi seems to yield overestimates.

    In the out-of-Africa model (fig \ref{fig:homsap-dfe.ooa}) the performance of the methods is more mixed. 
    In particular, variation in accuracy of the parameters estimated in each
    sample population depends on the methods used, and each method seems to 
    have its own biases with respect to the way in which population history 
    affects parameter estimates. For instance, for \dadi we see that estimates of 
    the shape parameter are most accurate in the YRI sample, while for \polydfe
    estimates are least accurate in the YRI sample. 
    With respect to the strength of selection, as in constant size populations,
    all methods are performing reasonably well, although \dadi seems to
    trail in terms of accuracy.

    %add PhoSin DFE figure here
    \begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/PhoSin_Vaquita2Epoch_1R22_Gamma_R22_Phocoena_sinus.mPhoSin1.pri.110_exons_DFE_plot.pdf}
\includegraphics{figures/PhoSin/Vaquita2Epoch_1R22/hs_hist.pdf}
        \caption{
        \label{fig:vaquita-dfe}
        A comparison of methods for inferring the distribution of fitness effects (DFE) from genetic data.
        Simulations were performed using a model of vaquita porpoise demography \citep{robinson2022critically} with a gamma-distributed DFE
        acting on exons parameterized by a mean selection coefficient, a shape parameter, and a relationship between selection and domiance. Estimates of the 
        parameters of the DFE are shown in the left ($\lvert E(s) \rvert $) and right hand panels (shape) respectively.
        This vaquita model has a single extant population, and parameter estimates from each
        of the three different methods are shown: 1) \grapes \cite{galtier2016adaptive}, 2) \polydfe \citep{tataru2020polydfe},
        and 3) \dadi \citep{gutenkunst2009inferring,kim2017inference}.
        C) Shown is the simulated distribution of $2 h s$, where $h$ is the dominance coefficient, along with the simulated distribution of $s$ and the median inferred DFEs. The distribution of $2 h s$ is multimodal because the simulated relationship between $h$ and $s$ is not continuous.
        % RNG: For C, need to extract inferred values from pipeline.
        }
    \end{figure}

    In Figures \ref{fig:vaquita-dfe.constant} and \ref{fig:vaquita-dfe} we show a comparison of estimates
    from the same three DFE-inference packages on simulations of the
    vaquita porpoise genome. In the constant size model (fig \ref{fig:vaquita-dfe.constant}) we see that all methods
    perform uniformly worse that in the human genome simulations, with consistent underestimation of the mean selection coefficient    
    and overestimation of the shape parameter. 
    A critical distinction is that the previous human simulations employed a DFE that assumed all selected mutations had additive effects
    (no dominance, $h = 1/2$), whereas the vaquita simulations use a DFE in which more deleterious mutations are more recessive ($h < 1/2$).
    For all three software packages, the inferred mean and shape of the distribution of selection coefficients 
    is consistent with the mean and shape of the distribution of $2 h s$, the product of the dominance coefficient
    and the selection coefficient (Fig.~\ref{fig:vaquita-dfe}C).
    This is likely because deleterious alleles are typically at low frequency and thus heterozygous, 
    where their selective effect is $h s$, and all these tools assume additivity  ($h = 1/2$).
    Strongly deleterious alleles are often observed to be recessive in many species \citep{mukai1972mutation},
    so these results suggest caution in interpreting DFE inferences.
    %RNG: Could use Kirk's help with the best references here

\section*{Detection of Selective Sweeps}
    \label{sweeps}
    We next highlight how the new models of selection in \stdpopsim can be used to benchmark and compare the performance of different methods for detecting selective sweeps from genetic data.
    Here, we performed simulations of human chromosome 1 under the three population out-of-Africa model from \cite{gutenkunst2009inferring} with purifying selection acting on exons under the gamma-distributed DFE described above, the HapMap II genetic map \citep{international2007second}, but also include sweeps.
    In particular, we introduced a beneficial mutation at a given location (to be varied) in the genome at a given time, and with a moderately strong selection coefficient ($s = 0.03$; $2Ns \sim 600$).
    For each sweep location, we performed 1000 replicate simulations, sampling from them only if the selected allele reached a frequency of 0.95 or greater.
    Rather than simulate the entire chromosome arm, replicate simulations were performed on 5cM regions of the chromosome at each of 100 different locations.
    As recombination rate varies across the genome, the size of the region in base pairs simulated varies, however the recombination distance, which is critical for the linked effects of a sweep, is held constant.

>>>>

% add commands for the DFE inference methods
\newcommand{\dfe}{\texttt{dfe-alpha}\xspace}
\newcommand{\polydfe}{\texttt{polyDFE}\xspace}
\newcommand{\dadicli}{\texttt{dadi-cli}\xspace}
\newcommand{\grapes}{\texttt{GRAPES}\xspace}

@@

\newcommand{\rngcomment}[1]{\textcolor{red}{RNG: #1}}

@@

% \author[1,+]{M. Elise Lauterbur}
% \author[3,*]{Ariella L. Gladstein}
 \author[4,*]{Graham Gower}
 \author[5*]{Nathaniel S. Pope}
 \author[5*]{Murillo F. Rodrigues}
 \author[5*]{Silas Tittes}
 \author[34*]{Linh N. Tran}
\author[6]{Georgia Tsambos}

\author[2]{Maria Izabel A. Cavassim}
% \author[5,7]{Jeff Adrion}
% \author[5]{Saurabh Belsare}
% \author[8]{Arjun Biddanda}
% \author[5]{Victoria Caudill}
% \author[9]{Jean Cury}
% \author[10]{Ignacio Echevarria}
% \author[11]{Benjamin C. Haller}
% \author[12,13]{Ahmed R. Hasan}
\author[14,15]{Xin Huang}
% \author[16]{Leonardo Nicola Martin Iasi}
% \author[17]{Ekaterina Noskova}
% \author[18]{Jana Obšteter}
% \author[19]{Vitor Antonio Corrêa Pavinato}
% \author[20,21]{Alice Pearson}
% \author[22,23]{David Peede}
% \author[24]{Manolo F. Perez}
\author[5]{Chris C. R. Smith}
% \author[25]{Jeffrey P. Spence}
% \author[5]{Anastasia Teterina}

@@

\affil[14]{\small{Department of Evolutionary Anthropology, University of Vienna, Vienna, Austria}}
\affil[15]{\small{Human Evolution and Archaeological Sciences (HEAS), University of Vienna, Vienna, Austria}}

@@

\affil[34]{\small{Department of Molecular and Cellular Biology, University of Arizona, Tucson AZ 85721, USA}}

@@

    % simulation
    To meet the growing need for the interpretation,
    analysis, and exploration of realistic and complex evolutionary models
    the field of population genetics has increasingly turned to simulation.
    Simulation in population genetics has a long history 
    including both backward-in-time coalescent simulations
    \citep{kingman1982genealogy,hudson1983testing, hudson1990gene}
    and forward-in-time simulations of complex demography and selection
    \citep[e.g.,][]{gillespie1984molecular,thornton2014c++, haller2019slim}.
    The development of simulation tools has been driven by the need to
    understand the effects of complex evolutionary processes on genetic
    variation \citep[e.g.,][]{galloway2020few}, to provide a null model for hypothesis testing
    \citep[e.g.,][]{hudson1992statistical,hudson1994evidence,sabeti2002detecting},
    to explore the power and limitations of statistical methods \citep[e.g.,][]{przeworski2002signature},
    and increasingly to provide a basis for machine learning and other
    simulation-based inference methods \citep[e.g.,][]{beaumont2002approximate,pavlidis2010searching,lin2011distinguishing,kern2018diplos,mughal2019localizing,sanchez2021deep,wang2021automatic}.
    While this is so, joint simulation of complex demography and selection
    is challenging,  requires a deep understanding of the underlying
    evolutionary processes, and necessitates a number of parameter choices including
    the strength of selection and the
    recombination rate, as well as a parameterized demographic model.
    This is a daunting task for many researchers, and can be a barrier to
    the adoption of simulation-based methods in population genetics.
    Furthermore, different simulation and inference tools may use different structures
    and parameter scalings for their models.

    % stdpopsim
    In light of the complexities associated with estimating and simulating
    population genetic models, it is no surprise that
    a lingering challenge with simulation in population genetics has been
    reproducibility and the ability to share and compare results among 
    researchers \citep[e.g.,][]{ragsdale2020lessons}.
    This challenge has been addressed in part by the development
    of community resources for sharing and distributing simulation software
    via the \stdpopsim project \citep{adrion2020community}. \stdpopsim
    provides a standardized interface for accessing a wide range of
    population genetic models, and has begun to be be widely adopted by the community. %maybe add citations here?
    While that is so, the original version of \stdpopsim did not include
    models of selection, which is a major limitation for more empirical
    applications of population genetic simulation. In particular, modeling selection
    through simulation is critical for understanding processes such
    as adaptation, the effects of selective sweeps \citep[e.g.][]{braverman1995hitchhiking,fay2000hitchhiking,przeworski2002signature,przeworski2005signature,schrider2015soft}, and the impact of
    background selection on genetic diversity
    \citep[e.g.][]{charlesworth1993effect,charlesworth1995pattern,williamson2002genealogy,ewing2016consequences,torres2020temporal}.
    Ideally, one would like
    to have a single, unified framework for simulating both neutral and
    non-neutral evolutionary processes, and to be able to compare the
    results of these simulations to empirical data in a manner that is
    both accessible to a wide range of researchers and highly reproducible. 
    Further the framework should include the complex realities of 
    genomes, including heterogenous recombination rates, 
    variation in the size and density of functional elements, and
    non-equilibrium demographic histories. Ideally, one should be able
    to model all of these features using estimates/annotations obtained
    from a species of interest \citep[e.g.][]{schrider2020background,rodrigues2024shared}.

@@

    \begin{wrapfigure}{l}{0.5\textwidth}
        \vspace{-0.0cm}
        \includegraphics[width=\linewidth]{figures/schematics/catalog.pdf}
        \caption{\label{fig:schematic}
        \textbf{(A)} A schematic of the \stdpopsim catalog and the different components that
        can be used in simulations. 
        \textbf{(B)} The python API and \textbf{(C)} the command line interface.
        The user specifies a species, a portion of
        the genome to simulate, and optionally a genetic map, a model of
        demography, and a model of selection that itself is composed of a
        distribution of fitness effects (DFE) and a set of functional
        annotations. \rngcomment{I find the images too small to read the text comfortably.}}
    \end{wrapfigure}
    \hfill

@@

    The vast majority of published DFE estimates are based on a limited number of species,
    so \stdpopsim currently has implemented published DFEs from
    four species: \textit{Arabidopsis thaleana} \citep{huber2018gene}, \textit{Drosophila melanogaster} \citep{ragsdale2016triallelic,huber2017determining},
    humans \citep{huber2017determining,kim2017inference}, and the vaquita porpoise \textit{Phocoena sinus} \citep{robinson2022critically}.
    While that is so, these DFEs can be applied to other species
    in the catalog. For instance, one could simulate a butterfly species with a gamma-distributed DFE originally
    estimated from humans, or a human population with a DFE estimated from \textit{Drosophila}.
    Furthermore, the user can specify a custom DFE and provide their own annotations
    of functional elements to simulate selection in a species for which we do not yet have 
    a published DFE included in the catalog. This flexibility allows for a wide range of
    models of selection to be simulated. %, and for the results of these simulations to be
    %compared to empirical data or used as a null model for hypothesis testing.
    % DRS: commenting that stuff out because there are obviously other use cases that could be
    % listed here (e.g. ML/ABC, and if we don't know enough about a species to do that then
    % we probably have no business doing hypothesis testing either, which could be just as
    % misleading or maybe even more so), so the simplest solution is to cut this as it is
    % not needed here.

@@

    Using \stdpopsim, we can easily simulate genetic data under a model of selection and then 
    examine the performance of different methods for inferring $N_e(t)$ from these simulated 
    data. In Figure \ref{fig:1pop-human-demography}, we show the results of such a benchmark
    where we have simulated genomes under a model of human out of Africa (OOA) demography
    with and without purifying selection acting on exons.
    For this simulation we have used
    a genetic map from the HapMap Project \citep{international2007second} along with a
    DFE for deleterious alleles (estimated by \cite{huber2017determining})
    that acts on exonic positions defined from the HAVANA group release 104 for the human genome.
    % DRS: Maybe we should say something about how easy it was to do these simulations, and maybe even show
    % some example code somewhere to prove our point? Could help emphasize the utility of this resource.
    These simulations were then used to estimate $N_e(t)$ using four methods: 1) \msmc \citep{Schiffels2020}, 
    2) \stairway \citep{liu2020stairway}, 3) \gone \citep{santiago2020recent}, and 4) \smcpp \citep{terhorst2017robust}.
    In this and what follows we performed 3 replicate whole genome simulations, so uncertainty in estimates might
    be roughly assessed.
    However we are mindful of wasted computational resources as thorough benchmarking
    is not the main purpose of this manuscript.

    In Figure \ref{fig:1pop-human-demography}, we show six panels, each of which shows the estimated $N_e(t)$
    from data simulated under the five-population model of human out-of-Africa demography
    from Ragsdale \textit{et al.} (2019)
    with samples taken from three modern geographic groups (CEU, CHB, and YRI from top to bottom).
    The left column shows the results of $N_e(t)$ estimates made from simulations without selection, 
    while the right column shows the results
    from simulations with deleterious mutations occurring in exons \rngcomment{(with inference based on simulated synonymous data from exons? or from whole genome?)}.
    In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$
    in various colors for the four methods. Dotted lines on the top row are to visually reinforce that inference is done
    from a model without selection. We see that these methods generally perform reasonably well in the absence of selection, 
    although GONE seems to overestimate $N_e(t)$ for the CHB sample and to a lesser extent in the CEU sample.
    Presumably this could be due to patterns of linkage disequilibrium that result from the combination
    of bottleneck and growth in these populations.
    In the presence of selection, different methods behave differently. 
    For instance stairwayplot, which uses the SFS to infer $N_e(t)$, 
    is reasonably close to the true $N_e(t)$ both with and without selection,
    although it seems to infer significant population size changes that are not part of the true history.
    Likewise \smcpp seems to be robust to the presence of selection,
    but the timing of population size changes is slightly inaccurate. 
    \msmc seems to be the most affected by the presence of selection, 
    estimates of $N_e(t)$ varying quite a bit from the true $N_e(t)$ particularly in the very recent past,
    although estimates from \msmc have quite a bit of added noise even in the absence of selection.

@@

    \begin{figure}[t]
        \centering
        \includegraphics[width=\textwidth]{figures/HomSap/OOA/estimated_Ne_t_final}
        \caption{
        \label{fig:1pop-human-demography}
        Performance of methods to infer $N_e(t)$ from a human out-of-Africa model \citep{ragsdale2019models}
        with and without purifying selection on exons. The left column shows estimates of $N_e(t)$ from simulations
        without selection, while the right column shows estimates of $N_e(t)$ from simulations with a gamma-distributed   
        DFE acting on exons. In each panel we show the true $N_e(t)$ in black, and the estimated $N_e(t)$ from four methods:    
        \msmc \citep{Schiffels2020}, \stairway \citep{liu2020stairway}, \gone \citep{santiago2020recent}, and \smcpp \citep{terhorst2017robust}.  
        \rngcomment{Need to update figure legend to remove ``all\_sites'' which seems not to be in plot.}
        \rngcomment{Why do some methods have confidence intervals (like stairwayplot) and others don't (like msmc)?}
        }
    \end{figure}
    %DRS-fig: reminder to update the legend to just say ``exons'' instead of ensembl_havana_whatever...

@@

\section*{Estimation of the Distribution of Fitness effects}
    \label{dfe}
    Another common application of population genetic inference is to estimate the distribution of fitness effects (DFE) of new mutations
    from genetic data. Here, we show how the new models of selection in \stdpopsim can be used to benchmark and compare
    the performance of different methods for inferring the DFE. We performed simulations using the 
    human and vaquita porpoise models described above, with complex demographic histories and models of selection acting on exons.

    A challenge when comparing quantitative DFE estimates among tools and with simulations is differing conventions of factors of two.
    The simulator SLiM and the inference tool \grapes define the selection coefficient $s$ such that a homozygote has fitness $1+s$.
    On the other hand, \dadicli and \polydfe define the selection coefficient such that a homozygote has fitness $1+ 2s$. %https://academic.oup.com/genetics/article/207/3/1103/5930676
    Moreover, all three inference tools directly infer the distribution of the selection coefficient scaled by an inferred ancestral effective population size $N_e$, but they assume different scaling.
    \dadicli infers the distribution of $2 N_e s$ while \polydfe and \grapes infer the distribution of $4 N_e s$. %https://academic.oup.com/genetics/article/207/3/1103/5930676, https://academic.oup.com/genetics/article/216/2/559/6066183
    For comparison, here we have normalized all inference results to the convention of SLiM.

    % Add citation to fastDFE as the successor to polyDFE: https://academic.oup.com/mbe/article/41/5/msae070/7641109 ?

    \begin{figure}[htbp]
        \centering
        \includegraphics[width=\linewidth]{figures/HomSap/OOA/HomSap_OutOfAfricaArchaicAdmixture_5R19_Gamma_K17_ensembl_havana_104_exons_DFE_plot}
        \caption{A comparison of methods for inferring the distribution of fitness effects (DFE) from genetic data.
        Simulations were performed using a human out-of-Africa model \citep{ragsdale2019models} with a gamma-distributed DFE
        acting on exons parameterized by a mean selection coefficient and a shape parameter. Estimates of the 
        parameters of the DFE are shown in the left ($\lvert E(s) \rvert $) and right hand panels (shape) respectively.
        The human model has three extant populations (CEU, CHB, and YRI), and parameter estimates from each
        sample are shown in different colors.
        Estimates from three different methods are shown: 1) \grapes \cite{galtier2016adaptive}, 2) \polydfe \citep{tataru2020polydfe},
        and 3) \dadicli \citep{Huang2023}.
            \rngcomment{Update figure legend. dadi=> dadi-cli.}
}
        \label{fig:homsap-dfe.ooa}
    \end{figure}

    In Figures \ref{fig:homsap-dfe.constant} and \ref{fig:homsap-dfe.ooa} we show a comparison of estimates
    from three software packages for inferring the DFE from genetic data:
    \dadicli \citep{Huang2023}, \polydfe \citep{tataru2020polydfe}, 
    and \grapes \citep{galtier2016adaptive}.
    In each case we note that the DFE is estimated from segregating sites only,
    without the use of substitution data, as a current limitation of \stdpopsim
    is that we are not able to simulate an outgroup sequence to estimate
    divergence. 
    Figure \ref{fig:homsap-dfe.constant} shows estimates from a single, constant size population of humans, 
    while Figure \ref{fig:homsap-dfe.ooa} shows estimates from a model of human out-of-Africa demography.
    In a constant size population, all methods perform reasonably well, although \dadicli seems to
    perform slightly worse than the other methods examined, at least for the small number of replicates we simulated here.
    Notably only \polydfe overlapped the true mean selection coefficient, whereas 
    the other methods do not. 
    For the shape parameter in constant size simulations \grapes and \polydfe are roughly
    equivalent in their performance whereas \dadicli seems to yield overestimates. 
    \rngcomment{Has Fig.~\ref{fig:homsap-dfe.constant} been updated for the new GRAPES scaling?}

    In the out-of-Africa model (Fig.~\ref{fig:homsap-dfe.ooa}) the performance of the methods is more mixed. 
    In particular, variation in accuracy of the parameters estimated in each
    sample population depends on the methods used, and each method seems to 
    have its own biases with respect to the way in which population history 
    affects parameter estimates. For instance, for \dadicli we see that estimates of 
    the shape parameter are most accurate in the YRI sample, while for \polydfe
    estimates are least accurate in the YRI sample. 
    With respect to the strength of selection, as in constant size populations,
    all methods are performing reasonably well, although \polydfe seems to
    lead in terms of accuracy.

    %add PhoSin DFE figure here
    \begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{figures/PhoSin/Vaquita2Epoch_1R22/PhoSin_Vaquita2Epoch_1R22_Gamma_R22_Phocoena_sinus.mPhoSin1.pri.110_exons_DFE_plot.pdf}
        \caption{
        \label{fig:vaquita-dfe}
        A comparison of methods for inferring the distribution of fitness effects (DFE) from genetic data.
        Simulations were performed using a model of vaquita porpoise demography \citep{robinson2022critically} with a gamma-distributed DFE
        acting on exons parameterized by a mean selection coefficient, a shape parameter, and a relationship between selection and dominance. Estimates of the 
        parameters of the DFE are shown in the left ($\lvert E(s) \rvert $) and right hand panels (shape) respectively.
        This vaquita model has a single extant population, and parameter estimates from each
        of the three different methods are shown: 1) \grapes \cite{galtier2016adaptive}, 2) \polydfe \citep{tataru2020polydfe},
        and 3) \dadicli \citep{Huang2023,kim2017inference}.
        C) Shown is the simulated distribution of $2 h s$, where $h$ is the dominance coefficient, along with the simulated distribution of $s$ and the median inferred DFEs. The distribution of $2 h s$ is multimodal because the simulated relationship between $h$ and $s$ is not continuous.
            \rngcomment{Update figure legend. dadi=> dadi-cli.}
        }
    \end{figure}

    In Figures \ref{fig:vaquita-dfe.constant} and \ref{fig:vaquita-dfe} we show a comparison of estimates
    from the same three DFE-inference packages on simulations of the
    vaquita porpoise genome. In both situations, we see that all methods
    perform uniformly worse that in the human genome simulations, with consistent underestimation of the mean selection coefficient    
    and overestimation of the shape parameter. 
    A critical distinction is that the previous human simulations employed a DFE that assumed all selected mutations had additive effects
    (no dominance, $h = 1/2$), whereas the vaquita simulations use a DFE in which more deleterious mutations are more recessive ($h < 1/2$).
    For all three software packages, the inferred mean and shape of the distribution of selection coefficients 
    is consistent with the mean and shape of the distribution of $2 h s$, the product of the dominance coefficient
    and the selection coefficient (Fig.~\ref{fig:vaquita-dfe}C).
    This is likely because deleterious alleles are typically at low frequency and thus heterozygous, 
    where their selective effect is $h s$, and all these tools assume additivity  ($h = 1/2$).
    Strongly deleterious alleles are often observed to be recessive in many species \citep{mukai1972mutation},
    so these results suggest caution in interpreting DFE inferences.
